---
title: "Collaborative working in git"
author: "K Purves"
date: 16.07.2020

output: rmdformats::material
---

# Introduction

This is an r notebook that goes along with a short workshop to get you working confidently and collaboratively using GitHub and local git repositories to ensure 1) version control and 2) regular code review.

You'll notice that this notebook has started off pretty empty and unimpressive. By the end of the workshop, we should have a nice completed project that we have built as a team, and we can confidently deploy in public. 

# Aims

By the end of the session, I hope that everyone in the room will have a github account, and some experience cloning existing repositories and working on documents collaboratively. 

## 1. Clone a project repository and link it with a remote

You should have started the workshop by cloning the remote (i.e. hosted on GitHub) repository that contains this code, and setting it up as a local git (i.e. saved on your personal machine or workspace).

Given cloning a repository initialises the folder as a git on your local machine,if you're reading this, you've already achieved aim 1. Congratulations. I hope that this fills you with determination. 

# Task 1 

## Objectives 

1. Learn how to read data in to r studio in a way that doesn't reveal the full download path in your script. This can be important if you are working on a shared server such as the LISA cluster where file paths cannot be shared between users.  
2. 

## Brief

We are going to play around with a dataset provided by fivethirtyeight that has a lot of variables on health and nutrition and some other variables. They used this dataset to demonstrate the danfgerts of p-hacking. 

[download the data here](https://data.fivethirtyeight.com/)

The data set is called nutrition-studies, from Jan 6 2016. Click the blue arrow to download it. Save it  


```{r}



```

this s a fun change again yay!!

# Task 2

# Task 3



# Analysis block

```{r  install packages}

#install.packages("rmdformats") # package with some great r notebook themes to play around with. Im testing some out.

```



