---
title: "Collaborative working in git"
author: "K Purves"
date: 16.07.2020

output: rmdformats::material
---

# Set up

install any packges that you might need
```{r  install packages}

#install.packages("rmdformats", "tidyverse") # package with some great r notebook themes to play around with. Im testing some out.

```


```{r Setup}
library(tidyverse)
```

# Introduction

This is an r notebook that goes along with a short workshop to get you working confidently and collaboratively using GitHub and local git repositories to ensure 1) version control and 2) regular code review.

You'll notice that this notebook has started off pretty empty and unimpressive. By the end of the workshop, we should have a nice completed project that we have built as a team, and we can confidently deploy in public. 

# Important points to come back to

A lot of these might not make sense at first, but I think it is useful to have them up front and here for reference. 

### Never work on the default (master) branch
The default (or master) should be code that is safe, reviewed and you're sure about. When youre adding new work, especially when many people might be doing things simultaneously, you might accidentally break someone elses code if you work directly on the default. 

### Start your session with the most up-to-date default branch
If you're going to work on a collaborative project, start your session by pulling the master branch. This will ensure that your days work starts with the most up to date version of the code.

### Commit often, push when your branch is ready to merge
A commit message makes it really clear what you have just done. So be sure to commit after every change or piece of work with a verbose message. When you push the code, you can start the process of raising a pull request and review, so it is usually best to do this when your project or piece of the project is ready to be reviewed.

### Do small pieces of work at a time, on a branch, then merge back to master
By this I mean try and make each branch a piece of work that fits well together, but is not huge. For example, creating the sumamry scores for the scales you are using in a porject might be done on one branch. Visualising the means of the summary scores would be another.


# Aims

By the end of the session, I hope that everyone in the room will have a github account, and some experience cloning existing repositories, working on documents collaboratively and reviewing each others code. 

## 1. Clone a project repository and link it with a remote

You should have started the workshop by cloning the remote (i.e. hosted on GitHub) repository that contains this code, and setting it up as a local git (i.e. saved on your personal machine or workspace).

Given cloning a repository initialises the folder as a git on your local machine,if you're reading this, you've already achieved aim 1. Congratulations. I hope that this fills you with determination. 

## 2. Run existing code and write your own on a new git branch

You should be able to run code you access from the remote on your local machine, and successfully add to it.

## 3. Push your new code to the remote git, and create a pull request

After adding code to this notebook, you will push it to GitHub and create a pull request to merge your new branch to the default branch.

## 4. Review and approve someone elses code, and merge it to the default branch

You should be able to use GitHub to identify a new pull request, review the code chunk, approve it and merge it to the default branch.

## 5. Identify and deal with a merge conflict

You should be able to identify and resolve a merge conflict, that results when the master branch changes while you are working on a branch. 


# Task 1 

## Objectives 

1. Learn how to read data in to r studio in a way that doesn't reveal the full download path in your script. This can be important if you are working on a shared server such as the LISA cluster where file paths cannot be shared between users.  
2. Run code that you have cloned from an existing repository on your local machine. 

## Brief

We are going to play around with a dataset provided by fivethirtyeight that has a lot of variables on health and nutrition and some other variables. They used this dataset to demonstrate the dangers of p-hacking. We'll use it because it has lots of variables in it so lots of room for people to play around with code without conflict.



## Activity 1

### 1.  The data set is called nutrition-studies, from Jan 6 2016. 

[download the data here](https://data.fivethirtyeight.com/)

Click the blue arrow to download it. Save it somewhere outside of your git folder, and unzip it.     

### 2.  Open the file called "File_info.R"
add the path to your new dataset (raw_anonymized_data.csv) , assigning it to the *datapath* variable.

### 3. Run the code chunk below. 
It will read in all variables assigned in the "File_info.R", including the path to your data. You should see 'datapath' appear as a variable in your environment 

**it is best practice to have this file on your local machine, but outside of the git so that it is not uploaded to GitHub or shared publically**

```{r read in source file}

source("./File_info.R")

```

*Note: there are other ways of ensuring path anonymity. For example, if you are working in an Rstudio project that is synced with git, and your data is in a folder just ouside of the git, you could use a datapath of ../data/dataname and this would also refrain from sharing the whole data path*

4. Read the data in

*either use option 1 as described in steps above, or unhash option 2 and ensure the data path is correct*

```{r read in datafile}

#option 1: read in path from source file

#df <- read_csv(datapath)

#option 2. Refer to folder just outside of project
df <- read_csv("../openData/nutrition-studies/raw_anonymized_data.csv")

```

5. View some basic details about the data

```{r view data}
# check data dimensions
dim(df)

# check data column names (first 50 to keep it manageable)
names(df[1:50])


```

# Task 2

## Objectives

1.  Create a new branch to work on   
2.  Write new code and commit it   

## Brief 

In this task, you will use the terminal to checkout a new branch, and add some small piece of code to this notebook. You will commit this code with a verbose commit message.

## Activity 2

### 1.  In your terminal, create and switch to a new branch by using the following command: **git checkout -b <branchname>**

Make sure your branch name is something distinctive, memorable and relavant for your task. Do not use spaces. An example might be "workshop_kirstin".

You do not want to name branches something generic, like "test" or "analysis" because you will run the risk of using the same branch name as someone else.

### 2.  In this notebook, add a chunk of code

It can be anything, but keep it simple and short. I suggest descriptive statistics for a variable of your choice, or cross tabs or correlation between two variables. 

Name your chunk something distinctive ( e.g. ```{r kirstin's first code chunk}```) and save your notebook.

### 3. Stage and commit your code

You now want to add your new changes to staging using the command **git add .** in your terminal

You then want to commit these changes to the local repository. Make sure you include a sensible commit message. The code for this is **git commit -m "your commit message describing what youve done"** and should again be used in the terminal. 

Make sure that you have navigated to the local git repository in your terminal before running any of this. 


# Task 3

## Objectives

1.  Push your code to the remote repository   
2.  Create a pull request for your branch

## Brief

This stage is key for collaborative work. You are now going to push the work you have done on your branch to the remote repository hosted on GitHub and create a pull request. 

A pull request is a signal to repository contributors that you think your code is ready to add to the default branch. It starts the process of code review and merging that is key to clear and reproducible code.

## Activities

### 1.  Push your committed code to the remote repository

The first step to pushing code from a new branch is to set the upstream repository / branch. You do this by telling git what the remote ("upstream") location is with the following command run in your terminal: **git --set-upstream https://github.com/klpurves/CollaborativeGitWorkshop.git <branchname>** 

Note that you need to use the branch name you created in task two at the end of this command.

If you weren't quite done, and needed to write more code and commit / push it before you go on to raise a pull request to finally merge the finished task, you could simply use **"git push"** on the command line for this step on this branch in the future. 

**Optional exercise**  
Add a second chunk of code on this branch, again give it a distinctive name. Stage (add), commit and push it before moving on to step 2.

### 2. Raise a pull request on github

Navigate to the repository on GitHub and raise a pull request. You might find slides 15-17 [of this presentation](https://docs.google.com/presentation/d/1v7zAdM7LgGWwgDtM3EVFMJd0AcKYqWYkXCJaB4AU-K4/edit?usp=sharing) a helpful reference for how to do this.

Be sure to explain what your code is meant to do, and any particular things you might want someone to check in the pull request message. 

Assign a reviewer to your pull request (see workshop slides).


# Task 4

## Objectives
1. Identify open pull requests and review code
2. Merge and delete branches

## Brief

This next task will all be done on GitHub. This stage of the process is crucial to checking and safely merging code, even if it may have been written simultaneously with a similar purpose on the same document. 

In this activity, you will identify open pull requests, complete a review, merge the successful code and delete the working branck. 

## Activites

### 1.  Identify an open pull request 

### 2.  View code, make comments and approve it

### 3.  Merge the new branch to the default

### 4.  Delete the approved and merged branch




